The Astropy Project

### 2017-10-17 main.tex
Update on Overleaf.

\documentclass[modern]{aastex61}

\newcommand{\escapecmd}[1]{\texttt{\detokenize{#1}}}

@@

% Packages / projects / programming - for consistency!
\newcommand{\package}[1]{\texttt{#1}}
\newcommand{\github}{\package{GitHub}}
\newcommand{\python}{\package{Python}}
\newcommand{\astropy}{Astropy}
\newcommand{\astropypkg}{\package{astropy}}

% For consistency:
\newcommand{\sectionname}{Section}
\renewcommand{\figurename}{Figure}
\newcommand{\equationname}{Equation}
\renewcommand{\tablename}{Table}

% For commenting - can be deleted before submission
\usepackage[colorinlistoftodos]{todonotes}
\newcommand{\inlinecomment}[2]{\todo[inline]{#1: #2}}
\newcommand{\comment}[2]{\todo{#1: #2}}

@@

\begin{abstract}
% I (Adrian) took a first stab at the abstract, but it needs some work. Feel
% free to modify!
The \astropy\ project supports and fosters the development of open-source
\python\ packages that provide commonly-needed functionality to the astronomical
community.
A key element of the project is the \astropy\ core package, which serves as the
foundation for more specialized projects and packages.
In this article, we provide an overview of the organization of the \astropy\
project and summarize key features in the core package as of the recent major
release, version 2.0.
We then describe the project infrastructure designed to facilitate and support
development for a broader ecosystem of interoperable packages.
We conclude with a future outlook of planned new features and directions for the
broader \astropy\ project.
\end{abstract}

@@

\begin{itemize}
   \item The goal is to produce a brief and informative paper that covers major
        \astropy\ principles not mentioned in the first paper, the core v2.0
        package, and infrastructure in \astropy\ project to support development
        in \python.
    \item We don't plan on including code in this paper, but if you think you
        will need to include code in your section, please add it to a separate
        Python module (.py file) and include it in this repository.
    \item Use \escapecmd{\sectionname} not ``Section,'' \escapecmd{\figurename}
        not ``Figure''
    \item Use \escapecmd{\astropypkg} for the astropy package,
        \escapecmd{\astropy} for the astropy project, \escapecmd{\python} not
        ``Python''
    \item If your subpackage was included in Paper I, then please just include a
        note on what the package does, a reference to paper I, and any new major
        updates to your package
    \item If your subpackage was not included, then please describe the
        sub-package on level with what was in the first paper, and highlight any
        major features in it. Typical length should be equivalent to one page
        (single-column).
    \item Please make sure you are logged into Overleaf or pushing a commit with
        your information to be able to track the contributors to the paper.

@@

\section{Introduction} \label{sec:intro}
% first draft by Moritz (hamogu)
All astronomical research makes use of software in some way and
astronomy as a field has long supported the development of specialized
software tools for different astronomical tasks. Some of those
software packages are (or at least were) supported by large
institutions, e.g.\ IRAF \citep[developed at NOAO,][]{IRAF}, MIDAS
\citep[developed at ESO,][]{MIDAS}, or ds9 \citep[developed at
% * <bsipocz@gmail.com> 2017-09-28T21:05:00.569Z:
%
% > or ds9
% probably starlink is a better example as being a large package of software similarly to IRAF and MIDAS rather than a single tool.
%
% ^.
% hamogu: Feel free ot swap in Starlink, if you think that's more appropriate. I picked ds9 becase (a) it's wider known (e.g. in the X-ray sounity) and (b) because it is an example of a widely used tool with institutional support that is NOT a large package. TOPCAT would have been another example if we want to expand that list.
  SAO,][]{ds9}. These packages tend to address a wide range of users
and they typically provide some level of documentation and user
support. Other packages are developed foremost by individual research
groups and used mostly in work by members of this group. Those
packages can address more specialized needs, but are not necessarily
available to the community and even if they are, documentation can in
some cases be insufficient for new users and the quality of the
software, i.e. the robustness and the level of testing that has been
performed to ensure correctness, is not always clear. On the other
hand, there are also many widely used, well documented, high quality
software packages which are developed by individuals or small groups.

The implementation of astronomical software, be it for a package meant
for wide distribution or scripts and programs for a specific research
project, can be eased by the use of a library that provides core
functionality that is common to many astronomical tasks. Reading and
writing FITS files is one example where it is time
consuming to implement the FITS standard correctly and
programming is much faster when using an existing library perform this
task. Another example of such a common task is dealing with different
astronomical coordinate systems. The \astropy\ project aims to provide a
library of core astronomical functionality in the \python\ programming
language.

@@

At the same time, an ecosystem of tools, programs, and webservices
has become available that makes it easier to collaborate on code development,
continuously test the code, build and provide documentation, and
supply downloadable packages. The \astropy\ project aims to develop and
provide high quality code and documentation according to the best
practices in software development. The project makes use of these
tools to do so without central institutional oversight.

The first public release of the \astropy\ package is described in
\cite{astropy}. Since then, the \astropy\ package has been
used in hundreds of projects, and the scope of the package as grown
considerably. At the same time, the community of astronomers
contributing to the project has grown tremendously and an ecosystem
of packages supporting or affiliated with the \astropy\ core has
developed. In this paper, we describe the current status of the
\astropy\ community, the \astropy\ core package, and the \astropy\
ecosystem.

We start by describing the way the \astropy\ project works and is
organized in \sectionname~\ref{sec:concepts}.  The project develops a core
package called \astropypkg\ (\sectionname~\ref{sec:core}) and several
separate packages that help setting up the infrastructure for testing,
documentation etc.\ (\sectionname~\ref{sec:infrastructure}). In
\sectionname~\ref{sec:learning} we give an overview of the wealth of
resources created to teach new users the use of the \astropy\ packages
and how to contribute to the community. We end with a short vision for
the future of \astropy\ in \sectionname~\ref{sec:future}.

@@

\subsection{APEs - Astropy Proposals for Enhancement}
%draft by hamogu
Central to the success of \astropy\ is an open environment where
anybody can contribute to the code by opening a so-called ``pull
request'' with a new code contribution on an publically available
website\footnote{\url{https://github.com/astropy/astropy}}; this
request is then reviewed by a maintainer responsible for the specific
sub-package and possibly merged into the main code base. Most bug
fixes and small features are contributed in this way. However, this
model leads to ``organic'' growth where different features are
implemented by different people with different programming styles and
interfaces. Thus, \astropy\ has a mechanism to more formally propose
changes to the code management (e.g.\ the release cycle) or to plan
out major new features. This mechanism is called ``Astropy Proposal
for Enhancement'' (APE). In an APE one or more authors lay out a
vision for a new feature, including a rationale why this feature is
needed, how it should be implemented, and what the interface will
be. These APEs are discussed and refined before much work is invested
into a detailed implementation. APEs are again edited and discussed on
a platform \footnote{\url{https://github.com/astropy/astropy-APEs}}
open to the public. In previous APEs a community concensus emerged and
APEs are accepted and become the basis for future work at this
point. In cases where consensus cannot be reached, the
\astropy\ coordination committee can decide to close the discussion and
make an executive decision based on the community input on the APE in
question. APEs are modeled after the ``Python Enhancement Proposals''
(PEP) and guide the development of the \python\ programming language.

@@

\subsection{Concept of affiliated packages}
% it may be described in previous subsection, if not the Brigitta can try to
% write it up
%
\par A major part of the \astropy{} project is the concept of “Affiliated
Packages”. An affiliated package is an astronomy-related \python{} package that
is not part of the \astropypkg{} core package, but has requested to be
included as part of the astropy project’s community. These packages support
the goals and vision of \astropy{} of improving code reuse,
interoperability, and embracing good coding practices such as testing and
thorough documentation. \astropypkg{} currently supports both versions of
\python\ 2.7+ and \python\ 3.4, affiliated packages are strongly encouraged to
support at least \python\ 3.x.
%
\par Affiliated packages contain functionality that is more specialized, or
have license incompatibility, or have enough external dependencies (e.g. GUI
libraries) that limit their integration possibility into the core and thus
they are more suitable to be separate packages. In general we hope that
becoming an affiliated package is seen as a good way for new and existing
packages to gain exposure.
%
% TODO: write about the process of becoming affiliated, and project packages
%
\subsection{Mixin}

@@

Historically, it has been unusual to keep units as part of numbers, especially as speed was often deemed essential and quantities used to be slower than regular arrays. Hence, programs and functions expected to be fed parameters in particular units, as (hopefully) stated in their documentation.  In \astropypkg\ 0.4, therefore, unit and quantity support was still somewhat spotty, but this has changed greatly.  One of the riskiest classes was angles, where both degrees and radians were ``obvious'' default units to some. As a consequence, quantities are most fully integrated in the coordinates module (\sectionname~\ref{sec:coordinates}). In \astropypkg\ 2.0, one can also opt in to use quantities fully in tables, by using the {\tt QTable} class (see \sectionname~\ref{sec:table}). As a result, units in FITS tables are also well-supported. For most other modules, quantities are at least accepted, and often assumed by default.

@@

\subsection{Release cycle and Long Term Support}
% This could also include information about the development cycle
% Brigitta can writes this up if there's no other taker
\inlinecomment{BMS: TODO}{reference APE2 for versioning and APE10 for dropping python2 support}
\inlinecomment{BMS}{shall we maybe refer to the release calendar, too?}
\inlinecomment{BMS}{should we mention that v3.0 is the exception to the
  version numbering scheme we use by not being 2.1?}
%
\par The general scheme for the releases and version numbering that the core
package has adopted if the following. \astropypkg\ has new feature releases every
6 months, and between feature releases additional bugfix releases that
contain only bugfixes but no new features. The timing of
%
\par Long-term support (LTS) releases continue to receive bugfixes for 2
years with no changes to the API. They are ideal for pipelines and other
applications where API stability is essential. The latest LTS release is
also the last one that supports \python 2, and will receive bug fixes until the
end of 2019.
%
\par The version numbering of \astropypkg\ reflects on the release scheme
described above. The core package uses the form x.y.z, where x is advances
on LTS, y is advanced for a non-LTS feature release, and z is advanced on
bugfix releases.

@@

The \astropy\ project aims to provide \python-based packages for all tasks that
are commonly needed in a large subset of the astronomical community.
At the foundation is the \astropypkg\ core package, which provides general
functionality (e.g., coordinate transformations) or base classes for other
packages to utilize for a common interface (e.g., \texttt{NDData}).
In this section, we highlight new features introduced or substantially improved
since version v0.2 (previously described in \citealt{astropy}).

@@

\subsubsection{Overview}
The \package{astropy.coordinates} subpackage is designed to support representing
and transforming celestial coordinates and, new in version 2.0, velocities.
The framework heavily relies on the \package{astropy.units} subpackage, and most
inputs to objects in this subpackage are expected to be \texttt{Quantity}
objects.
Some of the machinery also relies on the Essential Routines of Fundamental
Astronomy (ERFA) \texttt{C} library for some of the critical underlying
transformation machinery.
\inlinecomment{EJT}{Need to create a Zenodo DOI for ERFA and then cite it here.}
A key concept behind the design of this subpackage is that coordinate
\textit{representations} and \textit{reference systems / frames} are independent
of one another.
For example, a set of coordinates in the International Celestial Reference
System (ICRS) reference frame could be represented as spherical (right
ascension, declination, and distance from solar system barycenter) or Cartesian
coordinates ($x$, $y$, $z$ with the origin at ).
They can therefore change representations independent of being transformed to
other reference frames (e.g., the Galactic coordinate frame).

@@

\begin{description}
    \item[Local Earth coordinate frames] In addition to representing celestial
    coordinates, \astropypkg\ now supports specifying positions on the Earth in
    a number of different geocentric systems with the \texttt{EarthLocation}
    class.
    With this, \astropypkg\ now supports Earth-location-specific coordinate
    systems such as the altitude-azimuth (\texttt{AltAz}) or horizontal system.
    Transformations between \texttt{AltAz} and any Barycentric coordinate frame
    also requires specifying a time using the \texttt{Time} class from
    \texttt{astropy.time}.
    With this new functionality, many of the common tasks associated with
    observation planning can now be completed with \astropypkg\ or the
    \astropy-affiliated package \package{astroplan}. \inlinecomment{APW}{TODO: cite zenodo entry for astroplan?}.

@@

		The new \package{astropy.io.misc.yaml} module allows converting
		astropy objects into a standard YAML format.

@@

\astropypkg\ implements `normalized convolution' \citep[e.g.,][]{Knutsson1993}, which is an image reconstruction technique in which missing data are ignored during the convolution and replaced with values interpolated using the kernel.   In version $<=1.3$, the direct convolution and fft convolution approaches were not consistent, with fft convolution implementing normalized convolution and direct convolution implementing a different approach.  As of v2.0, the two methods are consistent and include a suite of consistency checks.

\begin{figure}
\includegraphics[width=\textwidth]{convolution_example.png}
An example showing different modes of convolution available in the \python\ ecosystem.  The red x's mark pixels that are set to NaN in the original data (a).  If the data are convolved with a Gaussian kernel on a 9x9 grid using scipy's direct convolution (b), any pixel within range of the original NaN pixels is also set to NaN.  Panel (c) shows what happens if the NaNs are set to zero first: the originally NaN regions are depressed relative to their surroundings.  Finally, panel (d) shows \astropypkg's convolution behavior, where the missing pixels are replaced with values interpolated from their surroundings using the convolution kernel.
\end{figure}

@@

The \astropypkg\ \package{stats} package provides
statistical tools that are useful for or specific to
astronomy and are not found in other \python\
statistics packages such as \package{scipy}
\citep{scipy} or \package{statsmodels}
\citep{seabold2010statsmodels}.   The
\package{stats} package contains a range of
different functionality used by many different
disciplines in astronomy.   It is not a complete set
of statistic tools, but rather a still growing
collection of useful features.

@@

\section{Infrastructure for affiliated packages}
\label{sec:infrastructure}
\inlinecomment{BMS}{should we use links to the appropriate github repo?}
\inlinecomment{BMS}{should we describe astropy-helpers even though it's future is unsure?}
%
\par In addition to astronomy related packages and libraries, the \astropy\
Project also contains and supports several infrastructure package that are
very generic. The following sections describes the most widely used
selection of them.
%
\subsection{Package template}
% Brigitta may write this up if there is no other takers
\par \astropy\ provides a template package that any \python\ package is welcome to
use (many affiliated packages do so). The template contains a ready-to-go
package layout. It provides infrastructure such as documentation tools,
testing framework and CI templates and configurations, Cython integration,
and documentation on how to make it all work.
%
\subsection{Continuous integration helpers}
% Brigitta
\package{ci-helpers} is a collection of scripts that empowers package
maintainers to control their testing set up and installation process for
various continuous integration services via a set of environment
variables. While the current development is mostly driven by the needs of
the \astropy\ ecosystem, the actual usage of this package is extremely
widespread. Currently we support setups for TravisCI and AppveyorCI.
%
\subsection{sphinx extensions}
% Tom R

@@

\section{The future of the Astropy project}
\label{sec:future}
\subsection{Roadmap for \python\ 3 only support}

@@

>>>>

\documentclass[modern]{aastex61}
\usepackage{xspace}
\newcommand{\escapecmd}[1]{\texttt{\detokenize{#1}}}

@@

% Packages / projects / programming - for consistency!
\newcommand{\package}[1]{\texttt{#1}\xspace}
\newcommand{\github}{\package{GitHub}\xspace}
\newcommand{\python}{\package{Python}\xspace}
\newcommand{\astropy}{Astropy\xspace}
\newcommand{\astropypkg}{\package{astropy}\xspace}

% For consistency:
\newcommand{\sectionname}{Section\xspace}
\renewcommand{\figurename}{Figure\xspace}
\newcommand{\equationname}{Equation\xspace}
\renewcommand{\tablename}{Table\xspace}

% For commenting - can be deleted before submission
\usepackage[colorinlistoftodos]{todonotes}
\newcommand{\inlinecomment}[2]{\todo[inline]{#1: #2}\xspace}
\newcommand{\comment}[2]{\todo{#1: #2}\xspace}

@@

\begin{abstract}
% I (Adrian) took a first stab at the abstract, but it needs some work. Feel
% free to modify!
The \astropy project supports and fosters the development of open-source
\python packages that provide commonly-needed functionality to the astronomical
community.
A key element of the project is the \astropy core package, which serves as the
foundation for more specialized projects and packages.
In this article, we provide an overview of the organization of the \astropy
project and summarize key features in the core package as of the recent major
release, version 2.0.
We then describe the project infrastructure designed to facilitate and support
development for a broader ecosystem of interoperable packages.
We conclude with a future outlook of planned new features and directions for the
broader \astropy project.
\end{abstract}

@@

\begin{itemize}
   \item The goal is to produce a brief and informative paper that covers major
        \astropy principles not mentioned in the first paper, the core v2.0
        package, and infrastructure in \astropy project to support development
        in \python.
    \item We don't plan on including code in this paper, but if you think you
        will need to include code in your section, please add it to a separate
        Python module (.py file) and include it in this repository.
    \item Use \escapecmd{\sectionname} not ``Section,'' \escapecmd{\figurename}
        not ``Figure''
    \item Use \escapecmd{\astropypkg} for the astropy package,
        \escapecmd{\astropy} for the astropy project, \escapecmd{\python} not
        ``Python''
    \item If your subpackage was included in Paper I, then please just include a
        note on what the package does, a reference to paper I, and any new major
        updates to your package
    \item If your subpackage was not included, then please describe the
        sub-package on level with what was in the first paper, and highlight any
        major features in it. Typical length should be equivalent to one page
        (single-column).
    \item Please make sure you are logged into Overleaf or pushing a commit with
        your information to be able to track the contributors to the paper.

@@

\section{Introduction} \label{sec:intro}
% first draft by Moritz (hamogu)
All astronomical research makes use of software in some way and
astronomy as a field has long supported the development of specialized
software tools for different astronomical tasks. Some of those
software packages are (or at least were) supported by large
institutions, e.g.~IRAF \citep[developed at NOAO,][]{IRAF}, MIDAS
\citep[developed at ESO,][]{MIDAS}, or ds9 \citep[developed at
% * <bsipocz@gmail.com> 2017-09-28T21:05:00.569Z:
%
% > or ds9
% probably starlink is a better example as being a large package of software similarly to IRAF and MIDAS rather than a single tool.
%
% ^.
% hamogu: Feel free ot swap in Starlink, if you think that's more appropriate. I picked ds9 becase (a) it's wider known (e.g. in the X-ray sounity) and (b) because it is an example of a widely used tool with institutional support that is NOT a large package. TOPCAT would have been another example if we want to expand that list.
  SAO,][]{ds9}. These packages tend to address a wide range of users
and they typically provide some level of documentation and user
support. Other packages are developed foremost by individual research
groups and used mostly in work by members of this group. Those
packages can address more specialized needs, but are not necessarily
available to the community and even if they are, documentation can in
some cases be insufficient for new users and the quality of the
software, i.e. the robustness and the level of testing that has been
performed to ensure correctness, is not always clear. On the other
hand, there are also many widely used, well documented, high quality
software packages which are developed by individuals or small groups.

The implementation of astronomical software, be it for a package meant
for wide distribution or scripts and programs for a specific research
project, can be eased by the use of a library that provides core
functionality that is common to many astronomical tasks. Reading and
writing FITS files is one example where it is time
consuming to implement the FITS standard correctly and
programming is much faster when using an existing library perform this
task. Another example of such a common task is dealing with different
astronomical coordinate systems. The \astropy project aims to provide a
library of core astronomical functionality in the \python programming
language.

@@

At the same time, an ecosystem of tools, programs, and webservices
has become available that makes it easier to collaborate on code development,
continuously test the code, build and provide documentation, and
supply downloadable packages. The \astropy project aims to develop and
provide high quality code and documentation according to the best
practices in software development. The project makes use of these
tools to do so without central institutional oversight.

The first public release of the \astropy package is described in
\cite{astropy}. Since then, the \astropy package has been
used in hundreds of projects, and the scope of the package as grown
considerably. At the same time, the community of astronomers
contributing to the project has grown tremendously and an ecosystem
of packages supporting or affiliated with the \astropy core has
developed. In this paper, we describe the current status of the
\astropy community, the \astropy core package, and the \astropy\
ecosystem.

We start by describing the way the \astropy project works and is
organized in \sectionname~\ref{sec:concepts}.  The project develops a core
package called \astropypkg (\sectionname~\ref{sec:core}) and several
separate packages that help setting up the infrastructure for testing,
documentation etc.~(\sectionname~\ref{sec:infrastructure}). In
\sectionname~\ref{sec:learning} we give an overview of the wealth of
resources created to teach new users the use of the \astropy packages
and how to contribute to the community. We end with a short vision for
the future of \astropy in \sectionname~\ref{sec:future}.

@@

\subsection{APEs - Astropy Proposals for Enhancement}
%draft by hamogu
Central to the success of \astropy is an open environment where
anybody can contribute to the code by opening a so-called ``pull
request'' with a new code contribution on an publically available
website\footnote{\url{https://github.com/astropy/astropy}}; this
request is then reviewed by a maintainer responsible for the specific
sub-package and possibly merged into the main code base. Most bug
fixes and small features are contributed in this way. However, this
model leads to ``organic'' growth where different features are
implemented by different people with different programming styles and
interfaces. Thus, \astropy has a mechanism to more formally propose
changes to the code management (e.g.~the release cycle) or to plan
out major new features. This mechanism is called ``Astropy Proposal
for Enhancement'' (APE). In an APE one or more authors lay out a
vision for a new feature, including a rationale why this feature is
needed, how it should be implemented, and what the interface will
be. These APEs are discussed and refined before much work is invested
into a detailed implementation. APEs are again edited and discussed on
a platform \footnote{\url{https://github.com/astropy/astropy-APEs}}
open to the public. In previous APEs a community concensus emerged and
APEs are accepted and become the basis for future work at this
point. In cases where consensus cannot be reached, the
\astropy coordination committee can decide to close the discussion and
make an executive decision based on the community input on the APE in
question. APEs are modeled after the ``Python Enhancement Proposals''
(PEP) and guide the development of the \python programming language.

@@

\subsection{Concept of affiliated packages}
% it may be described in previous subsection, if not the Brigitta can try to
% write it up
%
\par A major part of the \astropy project is the concept of “Affiliated
Packages”. An affiliated package is an astronomy-related \python package that
is not part of the \astropypkg core package, but has requested to be
included as part of the astropy project’s community. These packages support
the goals and vision of \astropy of improving code reuse,
interoperability, and embracing good coding practices such as testing and
thorough documentation. \astropypkg currently supports both versions of
\python~2.7 and \python~3.4+, affiliated packages are strongly encouraged to
support at least \python~3.x.
%
\par Affiliated packages contain functionality that is more specialized, or
have license incompatibility, or have enough external dependencies (e.g. GUI
libraries) that limit their integration possibility into the core and thus
they are more suitable to be separate packages. In general we hope that
becoming an affiliated package is seen as a good way for new and existing
packages to gain exposure.
%
% TODO: write about the process of becoming affiliated, and project packages
%
\subsection{Mixin}

@@

Historically, it has been unusual to keep units as part of numbers, especially as speed was often deemed essential and quantities used to be slower than regular arrays. Hence, programs and functions expected to be fed parameters in particular units, as (hopefully) stated in their documentation.  In \astropypkg~0.4, therefore, unit and quantity support was still somewhat spotty, but this has changed greatly.  One of the riskiest classes was angles, where both degrees and radians were ``obvious'' default units to some. As a consequence, quantities are most fully integrated in the coordinates module (\sectionname~\ref{sec:coordinates}). In \astropypkg~2.0, one can also opt in to use quantities fully in tables, by using the {\tt QTable} class (see \sectionname~\ref{sec:table}). As a result, units in FITS tables are also well-supported. For most other modules, quantities are at least accepted, and often assumed by default.

@@

\subsection{Release cycle and Long Term Support}
% This could also include information about the development cycle
% Brigitta can writes this up if there's no other taker
\inlinecomment{BMS: TODO}{reference APE2 for versioning and APE10 for dropping python2 support}
\inlinecomment{BMS}{shall we maybe refer to the release calendar, too?}
\inlinecomment{BMS}{should we mention that v3.0 is the exception to the
  version numbering scheme we use by not being 2.1?}
%
\par The general scheme for the releases and version numbering that the core
package has adopted if the following. \astropypkg has new feature releases every
6 months, and between feature releases additional bugfix releases that
contain only bugfixes but no new features. The timing of
%
\par Long-term support (LTS) releases continue to receive bugfixes for 2
years with no changes to the API. They are ideal for pipelines and other
applications where API stability is essential. The latest LTS release is
also the last one that supports \python 2, and will receive bug fixes until the
end of 2019.
%
\par The version numbering of \astropypkg reflects on the release scheme
described above. The core package uses the form x.y.z, where x is advances
on LTS, y is advanced for a non-LTS feature release, and z is advanced on
bugfix releases.

@@

The \astropy project aims to provide \python-based packages for all tasks that
are commonly needed in a large subset of the astronomical community.
At the foundation is the \astropypkg core package, which provides general
functionality (e.g., coordinate transformations) or base classes for other
packages to utilize for a common interface (e.g., \texttt{NDData}).
In this section, we highlight new features introduced or substantially improved
since version v0.2 (previously described in \citealt{astropy}).

@@

\subsubsection{Overview}
The \package{astropy.coordinates} subpackage is designed to support representing
and transforming celestial coordinates and, new in version 2.0, velocities.
The framework heavily relies on the \package{astropy.units} subpackage, and most
inputs to objects in this subpackage are expected to be \texttt{Quantity}
objects.
Some of the machinery also relies on the Essential Routines of Fundamental
Astronomy (ERFA) \texttt{C} library for some of the critical underlying
transformation machinery.
\inlinecomment{EJT}{Need to create a Zenodo DOI for ERFA and then cite it here.}
% * <tjenness@lsst.org> 2017-10-16T15:53:18.375Z:
% 
% I think it should mention that ERFA is related to SOFA (and cite SOFA).
% 
% ^.
A key concept behind the design of this subpackage is that coordinate
\textit{representations} and \textit{reference systems / frames} are independent
of one another.
For example, a set of coordinates in the International Celestial Reference
System (ICRS) reference frame could be represented as spherical (right
ascension, declination, and distance from solar system barycenter) or Cartesian
coordinates ($x$, $y$, $z$ with the origin at ).
They can therefore change representations independent of being transformed to
other reference frames (e.g., the Galactic coordinate frame).

@@

\begin{description}
    \item[Local Earth coordinate frames] In addition to representing celestial
    coordinates, \astropypkg now supports specifying positions on the Earth in
    a number of different geocentric systems with the \texttt{EarthLocation}
    class.
    With this, \astropypkg now supports Earth-location-specific coordinate
    systems such as the altitude-azimuth (\texttt{AltAz}) or horizontal system.
    Transformations between \texttt{AltAz} and any Barycentric coordinate frame
    also requires specifying a time using the \texttt{Time} class from
    \texttt{astropy.time}.
    With this new functionality, many of the common tasks associated with
    observation planning can now be completed with \astropypkg or the
    \astropy-affiliated package \package{astroplan}. \inlinecomment{APW}{TODO: cite zenodo entry for astroplan?}.

@@

		The new \package{astropy.io.misc.yaml} module allows converting
		astropy objects into a standard YAML format. 
        This can be used beyond \texttt{astropy.io.misc.yaml}: for example, to
        serialize the metadata of tables before saving to other formats like 
        HDF5.

@@

\astropypkg implements `normalized convolution' \citep[e.g.,][]{Knutsson1993}, which is an image reconstruction technique in which missing data are ignored during the convolution and replaced with values interpolated using the kernel.   In version $<=1.3$, the direct convolution and fft convolution approaches were not consistent, with fft convolution implementing normalized convolution and direct convolution implementing a different approach.  As of v2.0, the two methods are consistent and include a suite of consistency checks.

\begin{figure}
\includegraphics[width=\textwidth]{convolution_example.png}
An example showing different modes of convolution available in the \python ecosystem.  The red x's mark pixels that are set to NaN in the original data (a).  If the data are convolved with a Gaussian kernel on a 9x9 grid using scipy's direct convolution (b), any pixel within range of the original NaN pixels is also set to NaN.  Panel (c) shows what happens if the NaNs are set to zero first: the originally NaN regions are depressed relative to their surroundings.  Finally, panel (d) shows \astropypkg's convolution behavior, where the missing pixels are replaced with values interpolated from their surroundings using the convolution kernel.
\end{figure}

@@

The \astropypkg \package{stats} package provides
statistical tools that are useful for or specific to
astronomy and are not found in other \python\
statistics packages such as \package{scipy}
\citep{scipy} or \package{statsmodels}
\citep{seabold2010statsmodels}.   The
\package{stats} package contains a range of
different functionality used by many different
disciplines in astronomy.   It is not a complete set
of statistic tools, but rather a still growing
collection of useful features.

@@

\section{Infrastructure for affiliated packages}
\label{sec:infrastructure}
\inlinecomment{BMS}{should we use links to the appropriate github repo?}
\inlinecomment{BMS}{should we describe astropy-helpers even though it's future is unsure?}
%
\par In addition to astronomy related packages and libraries, the \astropy\
Project also contains and supports several infrastructure package that are
very generic. The following sections describes the most widely used
selection of them.
%
\subsection{Package template}
% Brigitta may write this up if there is no other takers
\par \astropy provides a template package that any \python package is welcome to
use (many affiliated packages do so). The template contains a ready-to-go
package layout. It provides infrastructure such as documentation tools,
testing framework and CI templates and configurations, Cython integration,
and documentation on how to make it all work.
%
\subsection{Continuous integration helpers}
% Brigitta
\package{ci-helpers} is a collection of scripts that empowers package
maintainers to control their testing set up and installation process for
various continuous integration services via a set of environment
variables. While the current development is mostly driven by the needs of
the \astropy ecosystem, the actual usage of this package is extremely
widespread. Currently we support setups for TravisCI and AppveyorCI.
%
\subsection{sphinx extensions}
% Tom R

@@

\section{The future of the Astropy project}
\label{sec:future}
\subsection{Roadmap for \python 3 only support}

@@

CI services? GitHub?
